---
title: 'Project 2nd attempt'
author: "Aashna"

date: "November 18, 2021"
output:
  html_document: default
  header-includes: \usepackage{color}
fontsize: 12pt
margin: 1in
---


```{r setup, echo=FALSE, message=F, warning=F}

    rm(list = ls())
#saving file to R files folder
setwd("/Users/aashnarungta/SCU/EconometricswithR/R_files/Project")

### Load the packages 
    library(doBy)
    library(dplyr)
    library(foreign)
    library(ggplot2)
    library(knitr)
    library(lmtest)
    library(readstata13)
    library(sandwich)
    library(stargazer)
    library(AER)
    library(gdata)
    library(wooldridge)
    library(openintro)
    library(tidyr)
    library(tidyverse)

cse=function(reg) {
    rob=sqrt(diag(vcovHC(reg, type="HC1")))
    return(rob)
    }
    
```



```{r data, echo=FALSE, message=F, warning=FALSE, include=FALSE}
#loading the data
student=read.csv("student-mat.csv", header=T, sep=",")
student$urban <- ifelse(student$address == 'U', 1, 0)
student$pstat <- ifelse(student$Pstatus == 'A', 1, 0)

```


```{r simple, echo=FALSE, message=F, warning=FALSE, comment=""}

#Making new column aggregating the 3 Grades columns
student <- student %>% rowwise() %>%
  mutate(Grades_total = sum(c_across(G1:G3)))

student$Grades <- with(student, (Grades_total / 60) * 100)



ggplot(student, aes(x=Dalc, y=Grades)) + geom_point(col="blue") +
        labs(title = "Daily Alc consumption on grades", x = "Daily Alcohol Consumption", y = "Grades", hjust=1, vjust=1) +
        stat_smooth(method=lm, col = "red", se=FALSE)


ggplot(student, aes(x=Walc, y=Grades)) + geom_point(col="blue") +
        labs(title = "Weekly Alc consumption on grades", x = "Weekly Alcohol Consumption", y = "Grades", hjust=1, vjust=1) +
        stat_smooth(method=lm, col = "red", se=FALSE)

```


```{r multiple, echo=FALSE, message=F, warning=FALSE, comment=""}

# Inserting scatterplot graphs of Grades on other variables.

reg1 = lm(Grades ~ Dalc, data = student)
reg2 = lm(Grades ~ Dalc + absences, data = student)
reg3 = lm(Grades ~ Dalc + absences + schoolsup , data = student)
reg4 = lm(Grades ~ Dalc + absences + schoolsup+ internet, data = student)
reg5 = lm(Grades ~ Dalc + absences + schoolsup+  internet + studytime, data = student)
reg6 = lm(Grades ~ Dalc + absences + schoolsup + internet + studytime + traveltime , data = student)

stargazer(reg1, reg2, reg3, reg4, reg5, reg6, se=list(cse(reg1), cse(reg2), cse(reg3), cse(reg4), cse(reg5), cse(reg6)), title="Grades against alcohol consumption", type="text", df=FALSE, digits=3)

```

Effect of alcohol consumption on grades:

We first run the regression model 1 with grades on daily alcohol consumption and see it is significant at 10% level. Since the model is a simple linear model, we suspect omitted variable bias and add other regressors to analyse the best baseline specification model.

As expected, we saw the fit of the model go up from 0.3% to 4.6% after adding other regressors. 

We want to study if alcohol consumption affects the grades of a student. From the star table, we notice there's a negative correlation but it is not statistically significant at standard confidence levels. 

From the regression table, we see absence from school too does not have a significant effect on the grades.
However, we notice the extra educational support has a very significant effect on grades with a 1% level.
Having internet also positively affects the grades with a significance level of 10%.
We noticed that study time positively affects the grades and is statistically significant at 5% whereas the travel time affects the grades negatively and is significant at 5% level.

To conclude, we reject our null hypothesis of alcohol consumption affecting grades. 

Since alcohol consumption does not explain the variation in grades, we wanted to take a look at another variable: the romantic relationship status of each student. We ran some regressions where our variable of interest was romantic, and the y-variable was grades.

```{r romantic, echo=FALSE, message=F, warning=FALSE, comment=""}

romantic1 = lm(Grades ~ romantic, data = student)
romantic2 = lm(Grades ~ romantic + studytime , data = student)
#romantic3 = lm(Grades ~ romantic + famrel + Pstatus, data = student)
romantic4 = lm(Grades ~ romantic + studytime+ goout, data = student)
romantic5 = lm(Grades ~ romantic + studytime + goout + age , data = student)
romantic6 = lm(Grades ~ romantic + studytime + goout + age +  address, data = student)
romantic7 = lm(Grades ~ romantic + studytime + goout + age + address + sex, data = student)


stargazer(romantic1, romantic2, romantic4, romantic5, romantic6, romantic7, se=list(cse(romantic1), cse(romantic2), cse(romantic4), cse(romantic5), cse(romantic6), cse(romantic7)), title="Grades against romantic lives", type="text", df=FALSE, digits=3)

```

In the above regressions, similar to the alcohol ones, we are looking at the grades of the same 395 students. We asked them about their relationship status, and created a dummy variable, where the variable takes a value of 1 if the student is in a romantic relationship, and 0 otherwise.

In the simple regression, it was seen that a student in a relationship will have, on average, lower grades by 4%. We decided to add some control variables to the regression in order to analyze the results and avoid any cases of omitted variable bias. The variables we added are:

1. Studytime - weekly study time. We added this as students in relationships are likely to be spending less time studying, and it is an important variable in determining school grades.

2. Goout - How often does the student go out with friends, on a scale of 1-5. This variable likely affects the likelihood of a student being in a relationship, and also dictates how much attention the student is paying to their grades.

3. Student age, from 15-22. This variable was important in determining how interested a student would be to have a romantic life.

4. Address - a binary variable that takes value of 1 if the student lives in an urban environment, and 0 if the student lives in a rural home. We suspect that a student who lives in a rural area will be less likely to be in a romantic relationship.

5. Sex - a binary variable that takes value of 1 if the student is male. We were curious to see if a student's sex affects their grades. 

From the above regressions, we can see that the romance variable has a negative relationship with grades - that is, students in a romantic relationship have grades that are lower by about 3%, keeping all other variables constant. However, we must keep in mind that this model only explains 8.2% of the variation in grades, and that the Romance variable is statistically significant at 10%. However, the adjusted $R^2$ is the highest in this regression, and therefore we chose for it to be our baseline. 

